---
title: "Untitled"
output: html_document
---
##Question 1
Split the data in two sets: A training set and a holdout set. We will use the first 39 weeks for training, the second 39 weeks for holdout testing (hint: The cut-off date is 30-09-1997).
```{r}
library(BTYD)
library(dplyr)
setwd("~/Documents/BISSmaster/SDSS/tut2")
load("CDnow data.RData")
cdnow <- rename(cdnow, date = Date, cust = ID, sales = N.CD, value = Value)
cdnow <- dc.MergeTransactionsOnSameDate(cdnow)#merge transactions
end.of.train.period <- as.Date("1997-09-30")
cdnow.train <- cdnow[which(cdnow$date <= end.of.train.period), ]
cdnow.test <- cdnow[which(cdnow$date > end.of.train.period), ]
```

##Question 2
```{r}
#To obtain the correct values requested in question 2, the BTYD package provides some convenient data cleaning (dc.) functions. Applying this set of functions results in the train.cbs object that contains all variables requested.
split.data <- dc.SplitUpElogForRepeatTrans(cdnow.train)#This function returns a filtered version where the first transaction is ignored because we are interested in repeat purchases
clean.cdnow <- split.data$repeat.trans.elog#This is the repeated purchase dataframe containing only the repeat purchases
freq.cbt <- dc.CreateFreqCBT(clean.cdnow) #This matrix is an overview of when repeat purchases for every customer occurred. The problem is though that the customers who did not do any repeat purchases are not in there, so we need to include them as well. This can be done as follows:
freq.cbt[1:3,1:3]
tot.cbt <- dc.CreateFreqCBT(cdnow)
train.cbt <- dc.MergeCustomers(tot.cbt, freq.cbt)
train.cbt[1:3,1:3]

#Now we can make a dataframe with the first purchase date, last purchase date and the end of the training period:
birth.periods <- split.data$cust.data$birth.per
last.dates <- split.data$cust.data$last.date
train.cbs.dates <- data.frame(birth.periods, last.dates, end.of.train.period)

#From the train.cbt and the train.cbs.dates, the train.cbs object can be built:
train.cbs <- dc.BuildCBSFromCBTAndDates(train.cbt, train.cbs.dates, per = "week")
train.cbs.df <- as.data.frame(train.cbs)
train.cbs.df[1:3,1:3]
#Here you can see the frequency of repeat purchases (x), the recency of the last purchase (t.x) and the total time that the customer was present (T.cal)
```

x: The number of repeat purchases (i.e. after the first purchase) for
customer i. This is 0 when no repeat purchases are made.
```{r}
#If we want to know the total number of repeat purchases per customer (where the customer ID is the index):
x <- as.data.frame(train.cbs)$x
```
T.x: The week of the last purchase. This can be a decimal number, e.g.
33.45 implies a purchase was made halfway through week 33.
```{r}
T.x <- as.data.frame(train.cbs)$t.x
```
T.cal: The length during which purchases could have occurred. This is the
time (in weeks) between the first purchase date and the cut-off date. The
maximum attainable value is 38.86 (i.e. one day after 01-01-1997, the
start of the data period)
```{r}
T.cal <- as.data.frame(train.cbs)$T.cal
```

##Question 3
Use the functions contained in the BTYD library. In particular, the function
bgnbd.EstimateParameters() allows you to estimate parameters for the
BG/NBD model. Make sure to pick different initial values for each
parameter, and avoid that b <1 as initial value.
```{r}
#to estimate the parameters we use the EstimateParameters() function:
params <- bgnbd.EstimateParameters(train.cbs)
#The log likelihood is:
LL <- bgnbd.cbs.LL(params, train.cbs)
#to make sure the model converges and is consistent, we run it a couple of times:
p.matrix <- c(params, LL)
for (i in 1:5){
  params <- bgnbd.EstimateParameters(train.cbs, params)
  LL <- bgnbd.cbs.LL(params, train.cbs)
  p.matrix.row <- c(params, LL)
  p.matrix <- rbind(p.matrix,p.matrix.row)
}
colnames(p.matrix)<- c("r", "alpha", "a", "b", "LL")
rownames(p.matrix)<- 1:6
p.matrix
#based on the result, it seem that the model converged, so it is ready to use.
params
```
##Question 4
Now we can create sales forecasts given the parameters. Use the
bgnbd.ExpectedCumulativeTransactions() to compute the cumulative number of
daily transactions for the remaining 39 weeks of the sample. Compare your
forecast to the actual number
```{r}
expectedTransactions <- bgnbd.ExpectedCumulativeTransactions(params, T.cal=train.cbs[,"T.cal"],T.tot=39, n.periods.final = 273)#with this function we can estimate how many transactions we'll receive by the end of the testing period. The actual number of transactions is by the end of that period is:
sprintf("The actual number of transactions in the test period is: %s",nrow(cdnow.test))
sprintf("The predicted number of transactions in the test period is: %s", tail(expectedTransactions, n=1))

#This shows that the model this shows that the model os off by 618 transactions
```

##Question 5
So far, we have not based our predictions on the past. Compute the expected
number of transactions in the final 39 weeks for a customer with history (x,
T.x,T.cal).
```{r}
x <- train.cbs["1516","x"]
t.x <- train.cbs["1516", "t.x"]
T.cal <- train.cbs["1516", "T.cal"]
bgnbd.ConditionalExpectedTransactions(params, T.star=52, x,t.x,T.cal)#Predictions for 52 weeks
```

##Extension
As future business analysts, we need to be able to advise the marketing department which customers to target and which customers to let go. In the Reinhartz and Kumar (2003) paper, a procedure is described to make these decisions. In the following case we'll implement the first 2 steps of this procedure, i.e. computing the NPV of the expected profit margin and comparing this to the marketing cost of a
If we need to make a decision on whether to spend advertising budget on customer 9, how would we go about making this decision? Use the paper from Reinhartz and Kumar (2003) as guideline, and the bgnbd.PAlive() function from the BTYD package. 

##Solution
To decide on whether we keep investing in customer 9 we need to do the following:
1. calculate the NPV of the expected contribution margin
2. compare this to the cost of marketing, and terminate if the marketing cost is higher than the NPV
3. calculate the finite lifetime estimates
4. conduct back-end performance analysis

We'll go through these steps step-by-step below:
```{r}
cdnow.train$rev <- cdnow.train$value #Compute revenue
cdnow.train$gross.profit <- (cdnow.train$rev*0.3) #compute gross profit
cdnow.train$weeknr <- as.numeric(sapply(cdnow.train$date, strftime, format = "%V")) #generate week number for each log entry
weeklycm <- tapply(cdnow.train$gross.profit, list(cdnow.train$cust, cdnow.train$weeknr), sum) #aggregate the gross profit for each customer to a weekly level
weeklycm[is.na(weeklycm)] <-0 #remove missing values
weeklycm <- (weeklycm-2)/7 #compute average weekly contribution margin
weeklycm <- rowSums(weeklycm)/39 #compute expected average weekly contribution margin

#Below we want to make a decision after 8 weeks whether or not to invest in the customer. To do this we discount the expected average weekly contribution margin for week 1 through 8 and multiply it with the probability of the customer being alive (as described in the paper and assuming an annual effective discount rate of 15%). Then, we compute the NPV of the marketing expenditures. If we assume that marketing costs amount to 2 euros per customer on a weekly basis we can compute the NPV of this expenditure flow. Then we examine whether the NPV of the expected contribution margin is larger than the NPV of the marketing expenditures. 
x <- train.cbs["9","x"]
t.x <- train.cbs["9", "t.x"]
T.cal <- train.cbs["9", "T.cal"]
weeklyDiscountFactor <- (1+0.15)^(1/52)

NPV_ECM <- c()
for(i in 1:8){
  discountedCM <- bgnbd.PAlive(params, x,t.x, T.cal) * weeklycm[9]*((1/(weeklyDiscountFactor))^i)
  NPV_ECM <- c(NPV_ECM, discountedCM)
}

marketingcostsTotal <- c()
for(i in 1:8){
  discountedMarketingCosts <- 2*((1/(weeklyDiscountFactor))^i)
  marketingcostsTotal <- c(marketingcostsTotal, discountedMarketingCosts)
}

train.cbs["9",]

sprintf("Should we spend advertising money on this customer?: %s because the NPV of ECM is %s and the discounted marketing costs are %s",ifelse(sum(NPV_ECM)>sum(marketingcostsTotal),"Yes","No"), sum(NPV_ECM),sum(marketingcostsTotal))
```



